---
title: "Multigroup Measurement Invariance"
author: "Elif Dede Yildirim"
date: "11/14/2022"
output: 
  html_document: 
    toc: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
bothsamples <- readRDS("H:/bothsamples.Rds")
```

```{r , include=FALSE}
library(car)
## Loading required package: carData
bothsamples$age <- recode(bothsamples$age, "-9=NA")
bothsamples$gender <- recode(bothsamples$gender, "-9=NA")
bothsamples$skill <- recode(bothsamples$skill, "-9=NA")
bothsamples$relax <- recode(bothsamples$relax, "-9=NA")
bothsamples$believe <- recode(bothsamples$believe, "-9=NA")
bothsamples$enjoy <- recode(bothsamples$enjoy, "-9=NA")
bothsamples$easy <- recode(bothsamples$easy, "-9=NA")
bothsamples$best <- recode(bothsamples$best, "-9=NA")
bothsamples$fun <- recode(bothsamples$fun, "-9=NA")
library(lavaan)

```

This R code sets up a multi-group confirmatory factor analysis (CFA) using the lavaan package. It defines the factor structure and relationships for two groups of data, and then estimates the model parameters and provides a summary of the results. The original models can be found in Bujacz et al. (2014) paper.  http://www.frontiersin.org/journal/10.3389/fpsyg.2014.00984/ 



# Measurement Invariance 

## Configural Invariance 

```{r}
# ; can be used to separate lines
# Define a multi-group CFA model
Configuralscale <- '
group: 1

#factor loadings for group 1
comfort =~ lambda11*relax +lambda12*easy
pleasure =~ lambda13*fun + lambda14*plea
content =~ lambda15*skill + lambda16*believe + lambda17*ideal + lambda18*best

#intercepts for group 1
relax~nu11*1; easy~nu12*1; fun~nu13*1; plea~nu14*1
skill~nu15*1; believe~nu16*1; ideal~nu17*1; best~nu18*1

#residual variances for group 1
relax~~theta11*relax; easy~~theta12*easy; fun~~theta13*fun; plea~~theta14*plea;
skill~~theta15*skill; believe~~theta16*believe; ideal~~theta17*ideal; best~~theta18*best

#factor means for group 1
comfort~0*1; pleasure~0*1; content~0*1

#factor variances for group 1
comfort~~1*comfort; pleasure~~1*pleasure; content~~1*content

#factor covariances for group 1
comfort~~psi11*pleasure; comfort~~psi12*content; pleasure~~psi13*content

group: 2
#factor loadings for group 2
comfort =~ lambda21*relax +lambda22*easy
pleasure =~ lambda23*fun + lambda24*plea
content =~ lambda25*skill + lambda26*believe + lambda27*ideal + lambda28*best

#intercepts for group 2
relax~nu21*1; easy~nu22*1; fun~nu23*1; plea~nu24*1
skill~nu25*1; believe~nu26*1; ideal~nu27*1; best~nu28*1

#residual variances for group 2
relax~~theta21*relax; easy~~theta22*easy; fun~~theta23*fun; plea~~theta24*plea;
skill~~theta25*skill; believe~~theta26*believe; ideal~~theta27*ideal; best~~theta28*best

#factor means for group 2
comfort~0*1; pleasure~0*1; content~0*1

#factor variances for group 2
comfort~~1*comfort; pleasure~~1*pleasure; content~~1*content

#factor covariances for group 2
comfort~~psi21*pleasure; comfort~~psi22*content; pleasure~~psi23*content
'
# Estimate the CFA model parameters 
Configuralscale.est <- lavaan(Configuralscale, data=bothsamples, group= "lang", missing="ML", auto.fix.first=FALSE)

# Display a summary of the estimated model, including fit measures, standardized parameter estimates, and R-squared values
summary(Configuralscale.est, fit.measures=TRUE, standardized=TRUE, rsq=T)


```


## Metric Invariance
```{r}

Metricscale <- '
group: 1
#factor loadings for group 1
comfort =~ lambda11*relax
comfort =~ lambda12*easy
pleasure =~ lambda13*fun
pleasure =~ lambda14*plea
content =~ lambda15*skill
content =~ lambda16*believe
content =~ lambda17*ideal
content =~ lambda18*best

#intercepts for group 1
relax ~ nu11*1
easy~ nu12*1
fun~ nu13*1
plea~ nu14*1
skill~ nu15*1
believe~ nu16*1
ideal~ nu17*1
best~ nu18*1

#residual variances for group 1
relax ~~ theta11*relax
easy~~ theta12*easy
fun~~ theta13*fun
plea~~theta14*plea
skill~~  theta15*skill
believe~~ theta16*believe
ideal~~  theta17*ideal
best~~  theta18*best

#factor means for group 1
comfort ~0*1
pleasure~0*1
content ~0*1

#factor variances for group 1
comfort ~~1*comfort
pleasure~~1*pleasure
content ~~1*content

#factor covariances for group 1
comfort ~~psi11*pleasure
comfort ~~psi12*content
pleasure~~psi13*content

group: 2
#factor loadings for group 2
comfort =~ lambda11*relax
comfort =~ lambda12*easy
pleasure =~ lambda13*fun
pleasure =~ lambda14*plea
content =~ lambda15*skill
content =~ lambda16*believe
content =~ lambda17*ideal
content =~ lambda18*best

#intercepts for group 2
relax ~ nu21*1
easy~ nu22*1
fun~ nu23*1
plea~ nu24*1
skill~ nu25*1
believe~ nu26*1
ideal~ nu27*1
best~ nu28*1

#residual variances for group 2
relax ~~ theta21*relax
easy~~ theta22*easy
fun~~ theta23*fun
plea~~theta24*plea
skill~~  theta25*skill
believe~~ theta26*believe
ideal~~  theta27*ideal
best~~  theta28*best

#factor means for group 2
comfort ~0*1
pleasure~0*1
content ~0*1

#factor variances for group 2
comfort ~~psi1*comfort
pleasure~~psi2*pleasure
content ~~psi3*content

#factor covariances for group 2
comfort ~~psi21*pleasure
comfort ~~psi22*content
pleasure~~psi23*content
'

# Estimate the CFA model parameters 

Metricscale.est <- lavaan(Metricscale, data=bothsamples, group= "lang", missing="ML", auto.fix.first=FALSE)

# Display a summary of the estimated model, including fit measures, standardized parameter estimates
summary(Metricscale.est, fit.measures=TRUE, standardized=TRUE)

```


## Model comparison
```{r}
lavTestLRT(Metricscale.est, Configuralscale.est )
```

## Partial Metric Invariance
```{r}
PMetricscale <- '
group: 1
#factor loadings for group 1
comfort =~ lambda11*relax
comfort =~ lambda12*easy
pleasure =~ lambda13*fun
pleasure =~ lambda14*plea
content =~ lambda15*skill
content =~ lambda16*believe
content =~ lambda17*ideal
content =~ lambda18*best

#intercepts for group 1
relax ~ nu11*1
easy~ nu12*1
fun~ nu13*1
plea~ nu14*1
skill~ nu15*1
believe~ nu16*1
ideal~ nu17*1
best~ nu18*1

#residual variances for group 1
relax ~~ theta11*relax
easy~~ theta12*easy
fun~~ theta13*fun
plea~~theta14*plea
skill~~  theta15*skill
believe~~ theta16*believe
ideal~~  theta17*ideal
best~~  theta18*best

#factor means for group 1
comfort ~0*1
pleasure~0*1
content ~0*1

#factor variances for group 1
comfort ~~1*comfort
pleasure~~1*pleasure
content ~~1*content

#factor covariances for group 1
comfort ~~psi11*pleasure
comfort ~~psi12*content
pleasure~~psi13*content

group: 2
#factor loadings for group 2
comfort =~ lambda21*relax
comfort =~ lambda12*easy
pleasure =~ lambda13*fun
pleasure =~ lambda14*plea
content =~ lambda15*skill
content =~ lambda16*believe
content =~ lambda17*ideal
content =~ lambda18*best

#intercepts for group 2
relax ~ nu21*1
easy~ nu22*1
fun~ nu23*1
plea~ nu24*1
skill~ nu25*1
believe~ nu26*1
ideal~ nu27*1
best~ nu28*1

#residual variances for group 2
relax ~~ theta21*relax
easy~~ theta22*easy
fun~~ theta23*fun
plea~~theta24*plea
skill~~  theta25*skill
believe~~ theta26*believe
ideal~~  theta27*ideal
best~~  theta28*best

#factor means for group 2
comfort ~0*1
pleasure~0*1
content ~0*1

#factor variances for group 2
comfort ~~psi1*comfort
pleasure~~psi2*pleasure
content ~~psi3*content

#factor covariances for group 2
comfort ~~psi21*pleasure
comfort ~~psi22*content
pleasure~~psi23*content
'
# Estimate the CFA model parameters 
PMetricscale.est <- lavaan(PMetricscale, data=bothsamples, group= "lang", missing="ML", auto.fix.first=FALSE)

# Display a summary of the estimated model, including fit measures, standardized parameter estimates, and R-squared values
summary(PMetricscale.est, fit.measures=TRUE, standardized=TRUE)

```


## Model comparison
```{r}
lavTestLRT(PMetricscale.est, Configuralscale.est)
```


## Strong Invariance
```{r}

Strongscale <- '
group: 1
#factor loadings for group 1
comfort =~ lambda11*relax
comfort =~ lambda12*easy
pleasure =~ lambda13*fun
pleasure =~ lambda14*plea
content =~ lambda15*skill
content =~ lambda16*believe
content =~ lambda17*ideal
content =~ lambda18*best

#intercepts for group 1
relax ~ nu11*1 
easy~ nu12*1
fun~ nu13*1
plea~ nu14*1
skill~ nu15*1
believe~ nu16*1
ideal~ nu17*1
best~ nu18*1

#residual variances for group 1
relax ~~ theta11*relax
easy~~ theta12*easy
fun~~ theta13*fun
plea~~theta14*plea
skill~~  theta15*skill
believe~~ theta16*believe
ideal~~  theta17*ideal
best~~  theta18*best

#factor means for group 1
comfort ~0*1
pleasure~0*1
content ~0*1

#factor variances for group 1
comfort ~~1*comfort
pleasure~~1*pleasure
content ~~1*content

#factor covariances for group 1
comfort ~~psi11*pleasure
comfort ~~psi12*content
pleasure~~psi13*content

group: 2
#factor loadings for group 2
comfort =~ lambda21*relax #free
comfort =~ lambda12*easy
pleasure =~ lambda13*fun
pleasure =~ lambda14*plea
content =~ lambda15*skill
content =~ lambda16*believe
content =~ lambda17*ideal
content =~ lambda18*best

#intercepts for group 2
relax ~ nu21*1 #free
easy~ nu12*1
fun~ nu13*1
plea~ nu14*1
skill~ nu15*1
believe~ nu16*1
ideal~ nu17*1
best~ nu18*1

#residual variances for group 2
relax ~~ theta21*relax
easy~~ theta22*easy
fun~~ theta23*fun
plea~~theta24*plea
skill~~  theta25*skill
believe~~ theta26*believe
ideal~~  theta27*ideal
best~~  theta28*best

#factor means for group 2
comfort ~alpha1*1
pleasure~alpha2*1
content ~alpha3*1

#factor variances for group 2
comfort ~~psi1*comfort
pleasure~~psi2*pleasure
content ~~psi3*content

#factor covariances for group 2
comfort ~~psi21*pleasure
comfort ~~psi22*content
pleasure~~psi23*content
'
# Estimate the CFA model parameters 
Strongscale.est<- lavaan(Strongscale, data=bothsamples, group= "lang", missing="ML", auto.fix.first=FALSE)

# Display a summary of the estimated model, including fit measures, standardized parameter estimates, and R-squared values
summary(Strongscale.est, fit.measures=TRUE, standardized=TRUE)

```

## Model comparison
```{r}
lavTestLRT(Strongscale.est, PMetricscale.est)
```



## Partial Strong Invariance
```{r}

PStrongscale <- '
group: 1
#factor loadings for group 1
comfort =~ lambda11*relax
comfort =~ lambda12*easy
pleasure =~ lambda13*fun
pleasure =~ lambda14*plea
content =~ lambda15*skill
content =~ lambda16*believe
content =~ lambda17*ideal
content =~ lambda18*best

#intercepts for group 1
relax ~ nu11*1
easy~ nu12*1
fun~ nu13*1
plea~ nu14*1
skill~ nu15*1
believe~ nu16*1
ideal~ nu17*1
best~ nu18*1

#residual variances for group 1
relax ~~ theta11*relax
easy~~ theta12*easy
fun~~ theta13*fun
plea~~theta14*plea
skill~~  theta15*skill
believe~~ theta16*believe
ideal~~  theta17*ideal
best~~  theta18*best

#factor means for group 1
comfort ~0*1
pleasure~0*1
content ~0*1

#factor variances for group 1
comfort ~~1*comfort
pleasure~~1*pleasure
content ~~1*content

#factor covariances for group 1
comfort ~~psi11*pleasure 
comfort ~~psi12*content
pleasure~~psi13*content

group: 2
#factor loadings for group 2
comfort =~ lambda21*relax #free
comfort =~ lambda12*easy
pleasure =~ lambda13*fun
pleasure =~ lambda14*plea
content =~ lambda15*skill
content =~ lambda16*believe
content =~ lambda17*ideal
content =~ lambda18*best

#intercepts for group 2
relax ~ nu21*1 #free
easy~ nu12*1
fun~ nu13*1
plea~ nu24*1 #free
skill~ nu25*1 #free
believe~ nu26*1 #free
ideal~ nu17*1
best~ nu18*1

#residual variances for group 2
relax ~~ theta21*relax
easy~~ theta22*easy
fun~~ theta23*fun
plea~~theta24*plea
skill~~  theta25*skill
believe~~ theta26*believe
ideal~~  theta27*ideal
best~~  theta28*best

#factor means for group 2
comfort ~alpha1*1
pleasure~alpha2*1
content ~alpha3*1

#factor variances for group 2
comfort ~~psi1*comfort
pleasure~~psi2*pleasure
content ~~psi3*content

#factor covariances for group 2
comfort ~~psi21*pleasure
comfort ~~psi22*content
pleasure~~psi23*content
'

# Estimate the CFA model parameters 
PStrongscale.est<- lavaan(PStrongscale, data=bothsamples, group= "lang", missing="ML", auto.fix.first=FALSE)

# Display a summary of the estimated model, including fit measures, standardized parameter estimates, and R-squared values
summary(PStrongscale.est, fit.measures=TRUE, standardized=TRUE)

```

## Model comparison
```{r}
lavTestLRT(PStrongscale.est,PMetricscale.est)
```

