---
title: "Latent Growth Modeling"
author: "EDY"
date: "11/15/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# Linear Growth Model
```{r}

plot_obs <- ggplot(data=mhealth_df, 
                   aes(x = Time, y = depression, group = id)) +       
                   geom_line() +                                        
                   theme_bw() +                                         
                   scale_x_continuous(breaks = 1:6, name = "Time") +   
                   scale_y_continuous(name = "Mental Health")      

#print the object (plot)
print(plot_obs)

# ex1: intercept-only growth model

icept <- lme(fixed = depression ~ 1 , random = ~ 1 |id, data = mhealth_df)
summary(icept)

# ex2: linear growth model

linear <- lme(fixed = depression ~ Time, random = ~ Time|id, data = mhealth_df)
summary(linear)


# ex3: quadratic growth model

quad <- lme(fixed = depression ~ Time + I(Time^2), random = ~ Time|id, data = mhealth_df)
summary(quad)

```

# No growth models
## No growth model using lme function.
```{r}

#fitting no growth model and assigning it to an object
depression.ng.lme <- lme(depression ~ 1, random= ~1 |id, data = mhealth_df, method="ML")

#obtaining summary of the model using the object we just created
summary(depression.ng.lme)
print(depression.ng.lme)
fixed.effects(depression.ng.lme)
```


## No growth model using "nlme" function.
```{r}

#fitting no growth model and assigning it to an object
depression.ng.nlme <- nlme(depression ~ beta_1 + d_1i,
                     data=mhealth_df,
                     fixed=beta_1~1,
                     random=d_1i~1,
                     group=~id,
                     start=c(beta_1=45),
                     na.action = "na.exclude")

#obtaining summary of the model using the object we just created
summary(depression.ng.nlme)
```

## No growth model using alternative specification in 'nlme'function.
```{r}

#fitting no growth model and assigning it to an object
depression.ng.nlme <- nlme(depression~b_1i,
                     data=mhealth_df,
                     fixed=b_1i~1,
                     random=b_1i~1|id,
                     start=c(b_1i=40),
                     na.action = "na.exclude")

#obtaining summary of the model using the object we just created
summary(depression.ng.nlme)

#No growth model using 'lme4' package,'lmer' function.
#fitting no growth model and assigning it to an object
depression.ng.lmer <- lmer(depression ~ 1 + (1 | id),
                     data = mhealth_df,
                     REML = FALSE,
                     na.action = "na.exclude")

#obtaining summary of the model using the object we just created
summary(depression.ng.lmer)
```

## No growth model using 'lavaan' package.

```{r}
names(mhealth_df)

# names of variables in data
#c("id","momed","Anx","Treatment","depression","Time")
mhealth_df_wide <- reshape(data=mhealth_df,
                          timevar=c("Time"),
                          idvar=c("id"),
                          v.names=c("depression","Anx","Treatment"),
                          direction="wide", sep="_")
#view the first few observations in the data set
head(mhealth_df_wide)
```

## no growth model as a lavaan model
```{r}
# no growth model in full SEM way
depression.ng.lavaan_model <- '
  # latent variable definitions
      #intercept
      eta_1 =~ 1*depression_1
      eta_1 =~ 1*depression_2
      eta_1 =~ 1*depression_4
      eta_1 =~ 1*depression_6
      

  # factor variances
      eta_1 ~~ eta_1

  # covariances among factors
      #none (only 1 factor)

  # factor means
      eta_1 ~ start(30)*1

  # manifest variances (made equivalent by naming theta)
      depression_1 ~~ theta*depression_1
      depression_2 ~~ theta*depression_2
      depression_4 ~~ theta*depression_4
      depression_6 ~~ theta*depression_6
     
  # manifest means (fixed at zero)
      depression_1 ~ 0*1
      depression_2 ~ 0*1
      depression_4 ~ 0*1
      depression_6 ~ 0*1
      
' 

#estimating the model using sem() function
depression.ng.lavaan_fit <- sem(depression.ng.lavaan_model,
                          data = mhealth_df_wide,
                          meanstructure = TRUE,
                          estimator = "ML",
                          missing = "fiml")

summary(depression.ng.lavaan_fit)

fitMeasures(depression.ng.lavaan_fit)

#Other summaries
parameterEstimates(depression.ng.lavaan_fit)
#inspect(depression.ng.lavaan_fit, what="est")
```

# Linear growth model using lme function.
```{r}

#fitting linear growth model and assigning it to an object
lg.depression.lme <- lme(depression ~ Time, random= ~ Time |id, data = mhealth_df, method="ML")

#obtaining summary of the model using the object we just created
summary(lg.depression.lme)

#Linear growth model using nlme function.
#fitting linear growth model and assigning it to an object
lg.depression.nlme <- nlme(depression~(beta_1+d_1i)+(beta_2+d_2i)*(Time),
                   data=mhealth_df,
                   fixed=beta_1+beta_2~1,
                   random=d_1i+d_2i~1,
                   group=~id,
                   start=c(beta_1=35,beta_2=4),
                   na.action = "na.exclude")

#obtaining summary of the model using the object we just created
summary (lg.depression.nlme)
```
#Linear growth model using alternative specification in nlme function.
```{r}

#fitting linear growth model and assigning it to an object
lg.depression.nlme <- nlme(depression~b_1i+b_2i*(Time),
                     data=mhealth_df,
                     fixed=b_1i+b_2i~1,
                     random=b_1i+b_2i~1|id,
                     start=c(b_1i=35, b_2i=4))

#obtaining summary of the model using the object we just created
summary(lg.depression.nlme)

#Linear growth model using lme4 package.
#fitting linear growth model and assigning it to an object
lg.depression.lmer <- lmer(depression ~ 1 + Time + (1 + Time | id), data = mhealth_df, REML = FALSE)

#obtaining summary of the model using the object we just created
summary(lg.depression.lmer)
```

## Linear growth model using lavaan package.
```{r}

#linear growth model
lg.depression.lavaan_model <- '
  # latent variable definitions
      #intercept (note intercept is a reserved term)
      eta_1 =~ 1*depression_1
      eta_1 =~ 1*depression_2
      eta_1 =~ 1*depression_4
      eta_1 =~ 1*depression_6
      

      #linear slope (note intercept is a reserved term)
      eta_2 =~ 0*depression_1
      eta_2 =~ 1*depression_2
      eta_2 =~ 3*depression_4
      eta_2 =~ 5*depression_6
      

  # factor variances
      eta_1 ~~ eta_1
      eta_2 ~~ eta_2

  # covariances among factors
      eta_1 ~~ eta_2

  # factor means
      eta_1 ~ start(35)*1
      eta_2 ~ start(4)*1

  # manifest variances (made equivalent by naming theta)
      depression_1 ~~ theta*depression_1
      depression_2 ~~ theta*depression_2
      depression_4 ~~ theta*depression_4
      depression_6 ~~ theta*depression_6
     
  # manifest means (fixed at zero)
      depression_1 ~ 0*1
      depression_2 ~ 0*1
      depression_4 ~ 0*1
      depression_6 ~ 0*1
      
' 


#estimating the model using sem() function
lg.depression.lavaan_fit <- sem(lg.depression.lavaan_model,
                          data = mhealth_df_wide,
                          meanstructure = TRUE,
                          estimator = "ML",
                          missing = "fiml")

summary(lg.depression.lavaan_fit)
```


```{r}
fitMeasures(lg.depression.lavaan_fit)

#Other summaries
parameterEstimates(depression.ng.lavaan_fit)
#inspect(depression.ng.lavaan_fit, what="est")
```
